Forecasting the Changing Dynamics of US Energy Product Exports to OECD Markets: The Role of the Russian-Ukraine Conflict and Future Scenarios

### 2023-05-19 report_Tomo editinta.Rmd
reikejo

```{r setup ETS, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

@@

```{r  comparison of ETS models}
#checking the accuracy
report(crude_fit)
report(additive_ets)
accuracy(crude_fit)
accuracy(additive_ets)

@@

Now, let's delve into another time series and examine the data pertaining to the reliance of OECD Europe countries on Russian crude oil exports. The goal for this overview is to check if our expectations to see opposite trend here compared to the US exports are correct. By analyzing this data, we will examine the autocorrelation function (ACF) to ensure that there is no presence of white noise. Additionally, we will utilize the ETS() function to generate a forecast, which we will then compare to the exports from the United States.
```{r  Overview of Reliance of OECD Europe on RUssian Crude Oil}
autoplot(oecd_relrus)
ACF(oecd_relrus) %>% autoplot()
```
Despite the relatively short time series period of only two years, the impact of the Russian invasion in Ukraine is evident in the graph, showcasing a significant decline in the reliance of OECD Europe countries on Russian crude oil. Additionally, after inspecting ACF graph, we can conclude that time series does not appear to be White Noise. Given this observation and the current situation, it is anticipated that this downward trend will likely persist in the future. Although the time series data exhibits some negative trends and errors, there is no apparent seasonality. Therefore, for this particular time series, we have chosen the ETS (M,A,N) model, taking into account the absence of seasonality and aiming to capture the underlying patterns and fluctuations.

@@

```{r Checking residual of ETS models, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#Plotting Residual
gg_tsresiduals(additive_ets)
gg_tsresiduals(auto_crude_ets)
gg_tsresiduals(relrus_fit)

@@

5.4 Comparison of ETS model against simple models
```{r Comparison of ETS model against simple models}

>>>>

```{r setup ETS, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

@@

```{r  comparison of ETS models, echo=FALSE}
#checking the accuracy
report(crude_fit)

report(additive_ets)

combined_accuracy1 <- 
  bind_rows(
    accuracy(additive_ets),
    accuracy(auto_crude_ets)
  )
combined_accuracy1

@@

Now, let's delve into another time series and examine the data pertaining to the reliance of OECD Europe countries on Russian crude oil exports. The goal for this overview is to check if our expectations to see opposite trend here compared to the US exports are correct. By analyzing this data, we will examine the autocorrelation function (ACF) to ensure that there is no presence of white noise. Additionally, we will utilize the ETS() function to generate a forecast, which we will then compare to the exports from the United States.
```{r  Overview of Reliance of OECD Europe on RUssian Crude Oil, echo=FALSE}
autoplot(oecd_relrus) + labs(title = "Reliance on Russian Crude Oil of OECD Europe")

ACF(oecd_relrus) %>% autoplot()+ labs(title = " ACF of Reliance on Russian Crude Oil of OECD Europe")
```
Despite the relatively short time series period of only two years, the impact of the Russian invasion in Ukraine is evident in the graph, showcasing a significant decline in the reliance of OECD Europe countries on Russian crude oil. Additionally, after inspecting ACF graph, we can conclude that time series does not appear to be White Noise. Given this observation and the current situation, it is anticipated that this downward trend will likely persist in the future. Although the time series data exhibits some negative trends and errors, there is no apparent seasonality. Therefore, for this particular time series, we have chosen the ETS (M,A,N) model, taking into account the absence of seasonality and aiming to capture the underlying patterns and fluctuations.

@@

```{r Checking residual of ETS models, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray',echo=FALSE}
#Plotting Residual
gg_tsresiduals(additive_ets)
gg_tsresiduals(auto_crude_ets)
gg_tsresiduals(relrus_fit)

@@

5.4 Comparison of ETS model against simple models
```{r Comparison of ETS model against simple models, echo=FALSE}

### 2023-05-19 report_Tomo-editinta.tex
reikejo

\DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
\CommentTok{\#Reliance on Russian gas}

\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{...}\DecValTok{1}\NormalTok{,}
               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Date\textquotesingle{}}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Total Oil Imports from Russia / Total Oil Imports (\%)\textquotesingle{}}\NormalTok{)}

\NormalTok{oecd\_relrus }\OtherTok{\textless{}{-}}\NormalTok{ ts\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(Country }\SpecialCharTok{\%in\%}\NormalTok{ oecd\_europe)}

@@

\NormalTok{total\_exp\_oecd }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_merged }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date, }\AttributeTok{names\_to =} \StringTok{"Export type"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"Amount"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{center}\includegraphics{report_files/figure-latex/Times series overivew-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/overview by product-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-2} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Seasonal plot for each product-3} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/seasonal-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/seasonal-2} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/seasonal-3} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-2} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Acf and white noise-3} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Plotting components-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Plotting components-2} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Plotting components-3} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Forecasting plots-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Forecasting plots-2} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}

@@

\begin{center}\includegraphics{report_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Combine the transformed time series}
\NormalTok{combined\_data }\OtherTok{\textless{}{-}} \FunctionTok{merge}\NormalTok{(crude\_oil\_bx, nat\_gas\_bx, }\AttributeTok{by =} \StringTok{"Date"}\NormalTok{, }\AttributeTok{all =} \ConstantTok{TRUE}\NormalTok{)}

\CommentTok{\#plotting time series}

\CommentTok{\#Determining the ETS Model for Crude OIL}
\FunctionTok{ggplot}\NormalTok{(combined\_data, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Date)) }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ Total\_Exports.x, }\AttributeTok{color =} \StringTok{"Crude Oil"}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ Total\_Exports.y, }\AttributeTok{color =} \StringTok{"Natural Gas"}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Amount (Transformed)"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Crude Oil and Natural Gas Exports"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_manual}\NormalTok{(}\AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\StringTok{"Crude Oil"} \OtherTok{=} \StringTok{"red"}\NormalTok{, }\StringTok{"Natural Gas"} \OtherTok{=} \StringTok{"green"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{report_files/figure-latex/setup ETS-1} \end{center}

@@

Within this section, we will explore various techniques and
methodologies, such as time series decomposition, selection of
appropriate forecasting methods, and evaluation of forecast accuracy.
The application of exponential smoothing and ETS models will enable us
to gain a deeper understanding of the patterns and dynamics within the
crude oil time series.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{addi\_fc}\OtherTok{\textless{}{-}}\NormalTok{ fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(additive\_ETS) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h=}\DecValTok{36}\NormalTok{)}

@@

\begin{center}\includegraphics{report_files/figure-latex/cars-1} \end{center}

@@

For the training dataset, the additive model had a lower RMSE (Root Mean
Squared Error) of 4.51 compared to 5.12 for the multiplicative model and
4.78 for the random model. The MAE (Mean Absolute Error) was also lower
for the additive model (3.52) compared to the multiplicative model
(3.99) and random model (3.65).

Additionally, the additive model had a lower MAPE (Mean Absolute
Percentage Error) of 7.37\% compared to 8.32\% for the multiplicative
model. The MASE (Mean Absolute Scaled Error) value for the additive
model (0.415) was lower than that of the multiplicative model (0.470),
indicating better forecasting accuracy.

Considering these metrics, the additive model demonstrated superior
performance in terms of forecasting accuracy for the natural gas time
series. Hence, we selected the additive model (ETS=AAA) as the most
appropriate ETS model for our analysis.

@@

After applying the ETS() function to choose an appropriate model for the
crude oil time series, we obtained the selected ETS model as
\textless ETS(A,Ad,N)\textgreater. To assess the accuracy of the
selected model, we calculated various metrics using the accuracy()
function. The results indicate that the chosen model performs reasonably
well in terms of forecasting accuracy. The ME (Mean Error) is close to
zero, indicating that, on average, the model's forecasts are unbiased.
The RMSE (Root Mean Squared Error) is 4.70, which represents the average
magnitude of the forecast errors. The MAE (Mean Absolute Error) is 3.64,
which measures the average absolute difference between the model's
forecasts and the actual values. The MPE (Mean Percentage Error) is
-1.33\%, indicating a slight underestimation of the crude oil exports.
The MAPE (Mean Absolute Percentage Error) is 7.87\%, providing a measure
of the relative forecasting accuracy. The MASE (Mean Absolute Scaled
Error) is 0.430, suggesting that the model's forecasts outperform a
naive seasonal method. The RMSSE (Root Mean Squared Scaled Error) is
0.464, which compares the model's performance to that of a naïve random
walk model. Finally, the ACF1 (Autocorrelation of Forecast Errors) is
0.183, representing the correlation between successive forecast errors.
Overall, the selected ETS model demonstrates satisfactory accuracy,
aligning with the main features observed in the crude oil time series.

@@

In contrast, the auto\_crude\_ets model (ETS(A,Ad,N)) chosen through
information criteria has different parameters and characteristics,
including smoothing parameters and initial states. Its information
criteria values are lower than those of the additive model.

@@

5.3 Checking the resuduals

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Plotting Residual}
\end{Highlighting}
\end{Shaded}

In context of white noise, in the case of the additive model and
crude\_fit, the significant values at lag 9 and lag 17 respectively
suggest that there is some correlation or pattern present in the
residuals at those specific lags.

Therefore, based on the ACF plot, it can be concluded that the residuals
in both models do not exhibit the characteristics of white noise. The
presence of significant values in the ACF plot indicates some level of
autocorrelation or dependency between the residuals at certain lags,
suggesting that there are systematic patterns or information that the
models have not fully captured. Yet to be started

@@

We start by performing Unit Root test to identify which methods are
suitable in our case
\includegraphics{report_files/figure-latex/Unit root test-1.pdf}

@@

\includegraphics{report_files/figure-latex/unnamed-chunk-1-1.pdf}
\includegraphics{report_files/figure-latex/unnamed-chunk-1-2.pdf}

>>>>

@@

\NormalTok{total\_exp\_oecd }\OtherTok{\textless{}{-}}\NormalTok{ oecd\_merged }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{Date, }\AttributeTok{names\_to =} \StringTok{"Export type"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"Amount"}\NormalTok{)}

\CommentTok{\#Reliance on Russian gas}

\FunctionTok{colnames}\NormalTok{(raw\_rel\_rusoil)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] "...1"       "01/01/2021" "01/02/2021" "01/03/2021" "01/04/2021"
##  [6] "01/05/2021" "01/06/2021" "01/07/2021" "01/08/2021" "01/09/2021"
## [11] "01/10/2021" "01/11/2021" "01/12/2021" "01/01/2022" "01/02/2022"
## [16] "01/03/2022" "01/04/2022" "01/05/2022" "01/06/2022" "01/07/2022"
## [21] "01/08/2022" "01/09/2022" "01/10/2022" "01/11/2022" "01/12/2022"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ raw\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{...}\DecValTok{1}\NormalTok{,}
               \AttributeTok{names\_to =} \StringTok{\textquotesingle{}Date\textquotesingle{}}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{\textquotesingle{}Total Oil Imports from Russia / Total Oil Imports (\%)\textquotesingle{}}\NormalTok{)}

\NormalTok{rel\_rusoil }\OtherTok{\textless{}{-}}\NormalTok{ rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}} \OtherTok{=} \FunctionTok{gsub}\NormalTok{(}\StringTok{"\%"}\NormalTok{, }\StringTok{""}\NormalTok{, }\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Date =} \FunctionTok{yearmonth}\NormalTok{(Date)) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{select}\NormalTok{(Date, Country, }\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}}\NormalTok{)}

\NormalTok{oecd\_relrus }\OtherTok{\textless{}{-}}\NormalTok{ ts\_rel\_rusoil }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(Country }\SpecialCharTok{==} \StringTok{"OECD Europe"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}} \OtherTok{=} \FunctionTok{as.numeric}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}}\NormalTok{))  }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{drop\_na}\NormalTok{(Date) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{drop\_na}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{Total Oil Imports from Russia / Total Oil Imports (\%)}\StringTok{\textasciigrave{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Times series overivew-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/overview by product-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Seasonal plot for each product-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Seasonal plot for each product-2} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Seasonal plot for each product-3} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/seasonal-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/seasonal-2} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/seasonal-3} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Acf and white noise-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Acf and white noise-2} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Acf and white noise-3} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Plotting components-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Plotting components-2} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Plotting components-3} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Forecasting plots-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Forecasting plots-2} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Residual diagnostics (crude oil)-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Residual diagnostics (crude oil)-2} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Residual diagnostics (nat gas)-1} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Residual diagnostics (nat gas)-2} \end{center}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/setup ETS-1} \end{center}

@@

Within this section, we will explore various techniques and
methodologies, such as selection of appropriate forecasting methods, and
evaluation of forecast accuracy. The application of exponential
smoothing and ETS models will enable us to gain a deeper understanding
of the patterns and dynamics within the crude oil time series.

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{report}\NormalTok{(fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 3 x 9
##   .model              sigma2 log_lik   AIC  AICc   BIC   MSE  AMSE   MAE
##   <chr>                <dbl>   <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
## 1 additive_ETS       84.0      -463.  960.  968. 1005.  71.1  78.4 6.45 
## 2 multiplicative_ETS  0.0351   -500. 1033. 1040. 1078. 106.  113.  0.116
## 3 random1_ETS         0.0321   -495. 1023. 1030. 1068.  74.9  81.8 0.104
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{additive\_ets}\OtherTok{\textless{}{-}}\NormalTok{fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(additive\_ETS)}

\NormalTok{addi\_fc}\OtherTok{\textless{}{-}}\NormalTok{ fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(additive\_ETS) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h=}\DecValTok{36}\NormalTok{)}

@@

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Determing ETS model for Crude Oil exports-1} \end{center}

@@

For the training dataset, the additive model had the lowest ME (Mean
Error) of -0.036, suggesting that it has a relatively small average
difference between its predictions and the actual values. Also ACF1 only
of 0.11, which suggests that the residuals are reasonably independent
and do not exhibit strong patterns or dependencies.

Additionally, the additive model had the lowest AICc, AIS and BIC
compared to two other models, indicating that it provides a better
balance between model complexity and goodness of fit. The lower values
of AICc, AIC, and BIC suggest that the additive model outperforms the
other models in terms of explaining the data while considering the
number of parameters. This implies that the additive model is a more
parsimonious choice and provides a better fit to the observed data,
making it a preferred option for this analysis.

Considering these metrics, the additive model demonstrated superior
performance in terms of forecasting accuracy for the natural gas time
series. Hence, we selected the additive model (ETS=AAA) as the most
appropriate ETS model for our analysis. Ultimately, upon plotting the
forecast, it becomes evident that the additive model predicts a
continuation of the upward trend in exports.

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#ETS model selection by R}
\NormalTok{crude\_fit }\OtherTok{\textless{}{-}}\NormalTok{ crude\_oil\_bx }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{model}\NormalTok{(}\AttributeTok{auto\_crude\_ets =} \FunctionTok{ETS}\NormalTok{(Total\_Exports))}

\FunctionTok{accuracy}\NormalTok{(crude\_fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 1 x 10
##   .model         .type         ME  RMSE   MAE   MPE  MAPE  MASE RMSSE  ACF1
##   <chr>          <chr>      <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
## 1 auto_crude_ets Training -0.0372  8.75  6.66  -Inf   Inf 0.432 0.462 0.130
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{auto\_crude\_ets}\OtherTok{\textless{}{-}}\NormalTok{crude\_fit }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{select}\NormalTok{(auto\_crude\_ets)}
\NormalTok{auto\_crud}\OtherTok{\textless{}{-}}\NormalTok{auto\_crude\_ets }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{forecast}\NormalTok{(}\AttributeTok{h=}\DecValTok{36}\NormalTok{)}

\NormalTok{auto\_crud}\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{autoplot}\NormalTok{(}
\NormalTok{    crude\_oil\_bx) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{y =} \StringTok{"Thousand Barrels"}\NormalTok{,}
    \AttributeTok{title =} \StringTok{"Forecasts for U.S. Crude oil exports to OECD Europe"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{guides}\NormalTok{(}\AttributeTok{colour =} \FunctionTok{guide\_legend}\NormalTok{(}\AttributeTok{title =} \StringTok{"Forecast"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/applying ETS()-1} \end{center}

After applying the ETS() function to choose an appropriate model for the
crude oil time series, we obtained the selected ETS model as
\textless ETS(A,A,N)\textgreater. To assess the accuracy of the selected
model, we calculated various metrics using the accuracy() function. The
results indicate that the chosen model performs reasonably well in terms
of forecasting accuracy. The ME (Mean Error) is close to zero,
indicating that, on average, the model's forecasts are unbiased. The ME
of the model is -0.037 which is relatively small. Also, the ACF1
(Autocorrelation of Forecast Errors) is 0.129, representing the
correlation between successive forecast errors. Finally, AIC, AICc and
BIC cafeterias are also small indicating that model is a good fit in
this case. Overall, the selected ETS model demonstrates satisfactory
accuracy, aligning with the main features observed in the crude oil time
series.

\begin{verbatim}
## Series: Total_Exports 
## Model: ETS(A,A,N) 
##   Smoothing parameters:
##     alpha = 0.2625271 
##     beta  = 0.000100007 
## 
##   Initial states:
##      l[0]     b[0]
##  19.51402 1.034277
## 
##   sigma^2:  79.5955
## 
##      AIC     AICc      BIC 
## 944.1412 944.7535 957.3632
\end{verbatim}

\begin{verbatim}
## Series: Total_Exports 
## Model: ETS(A,A,A) 
##   Smoothing parameters:
##     alpha = 0.2721743 
##     beta  = 0.0001004441 
##     gamma = 0.0001002519 
## 
##   Initial states:
##      l[0]     b[0]       s[0]    s[-1]    s[-2]      s[-3]     s[-4]    s[-5]
##  18.14101 1.043379 -0.3552452 1.289356 2.580988 -0.4585203 0.4696732 1.414157
##      s[-6]    s[-7]   s[-8]     s[-9]    s[-10]    s[-11]
##  -1.414565 3.153327 1.55624 0.2737964 -3.786328 -4.722879
## 
##   sigma^2:  84.0217
## 
##       AIC      AICc       BIC 
##  960.4748  967.5911 1005.4294
\end{verbatim}

\begin{verbatim}
## # A tibble: 2 x 10
##   .model         .type         ME  RMSE   MAE   MPE  MAPE  MASE RMSSE  ACF1
##   <chr>          <chr>      <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
## 1 additive_ETS   Training -0.0360  8.43  6.45  -Inf   Inf 0.418 0.445 0.114
## 2 auto_crude_ets Training -0.0372  8.75  6.66  -Inf   Inf 0.432 0.462 0.130
\end{verbatim}

@@

In contrast, the auto\_crude\_ets model (ETS(A,A,N)) chosen through
information criteria has different parameters and characteristics,
including smoothing parameters and initial states. Its information
criteria values are lower than those of the additive model.

@@

Now, let's delve into another time series and examine the data
pertaining to the reliance of OECD Europe countries on Russian crude oil
exports. The goal for this overview is to check if our expectations to
see opposite trend here compared to the US exports are correct. By
analyzing this data, we will examine the autocorrelation function (ACF)
to ensure that there is no presence of white noise. Additionally, we
will utilize the ETS() function to generate a forecast, which we will
then compare to the exports from the United States.

\begin{verbatim}
## Plot variable not specified, automatically selected `.vars = Total Oil Imports
## from Russia / Total Oil Imports (%)`
\end{verbatim}

\includegraphics{report_Tomo-editinta_files/figure-latex/Overview of Reliance of OECD Europe on RUssian Crude Oil-1.pdf}

\begin{verbatim}
## Response variable not specified, automatically selected `var = Total Oil
## Imports from Russia / Total Oil Imports (%)`
\end{verbatim}

\includegraphics{report_Tomo-editinta_files/figure-latex/Overview of Reliance of OECD Europe on RUssian Crude Oil-2.pdf}
Despite the relatively short time series period of only two years, the
impact of the Russian invasion in Ukraine is evident in the graph,
showcasing a significant decline in the reliance of OECD Europe
countries on Russian crude oil. Additionally, after inspecting ACF
graph, we can conclude that time series does not appear to be White
Noise. Given this observation and the current situation, it is
anticipated that this downward trend will likely persist in the future.
Although the time series data exhibits some negative trends and errors,
there is no apparent seasonality. Therefore, for this particular time
series, we have chosen the ETS (M,A,N) model, taking into account the
absence of seasonality and aiming to capture the underlying patterns and
fluctuations.

\begin{verbatim}
## # A tibble: 1 x 11
##   Country     .model     .type     ME  RMSE   MAE   MPE  MAPE  MASE RMSSE   ACF1
##   <chr>       <chr>      <chr>  <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>
## 1 OECD Europe relrus_ETS Trai~ -0.223  1.98  1.48 -2.16  8.13 0.272 0.319 0.0151
\end{verbatim}

\includegraphics{report_Tomo-editinta_files/figure-latex/Forecast for reliance-1.pdf}
Upon applying the ETS model to the reliance time series, we observe a
satisfactory level of accuracy, as indicated by the low MA and ACF1
values. Subsequently, we generated a forecast for the upcoming year,
revealing that according to our model, the reliance is expected to
continue decreasing. However, it is important to note that this forecast
is based solely on two years of data, which may limit its accuracy and
reliability.

Upon examining the forecast presented in this chapter, it becomes
evident that in response to the Russian invasion, OECD Europe
significantly reduced its reliance on Russian crude oil and began
increasing imports from the United States. Our forecast indicates that
this trend of decreased reliance on Russian oil and increased imports
from the US is expected to persist in the future.

5.3 Checking the residuals

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Checking residual of ETS models-1} \end{center}

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Checking residual of ETS models-2} \end{center}

\begin{center}\includegraphics{report_Tomo-editinta_files/figure-latex/Checking residual of ETS models-3} \end{center}

After checking the residuals of our selected ETS model for Crude oil
exports and Reliance on Russian crude oil as well as the one selected
automatically by R for exports, we can see that in all cases residual
are normally distributed around zero which implies that there is no
correlation between the residuals and thus suggesting that the models
are accurate in the predictions of the data.

5.4 Comparison of ETS model against simple models

\begin{verbatim}
## # A tibble: 6 x 10
##   .model         .type     RMSE   MAE  MAPE  MASE RMSSE      ME   MPE   ACF1
##   <chr>          <chr>    <dbl> <dbl> <dbl> <dbl> <dbl>   <dbl> <dbl>  <dbl>
## 1 Drift          Training 11.3   8.63 Inf   0.509 0.538 NA         NA NA    
## 2 Mean           Training 30.9  27.4  Inf   1.62  1.47  NA         NA NA    
## 3 Naive          Training 11.4   8.68 Inf   0.512 0.540 NA         NA NA    
## 4 SNaive         Training 21.0  17.0   24.7 1     1     NA         NA NA    
## 5 additive_ETS   Training  8.43  6.45 Inf   0.418 0.445 -0.0360  -Inf  0.114
## 6 auto_crude_ets Training  8.75  6.66 Inf   0.432 0.462 -0.0372  -Inf  0.130
\end{verbatim}

In conclusion, when comparing our ETS models to the simple models from
the Toolbox chapter, we consistently observe superior performance by the
ETS models across all accuracy measures. This highlights the significant
enhancement in accuracy achieved through the Exponential Smoothing
approach. The ETS models consistently outperform the simple models,
affirming the effectiveness and value of employing Exponential Smoothing
techniques in forecasting.

@@

We start by performing Unit Root test to identify which methods are
suitable in our case
\includegraphics{report_Tomo-editinta_files/figure-latex/Unit root test-1.pdf}

@@

\includegraphics{report_Tomo-editinta_files/figure-latex/unnamed-chunk-1-1.pdf}
\includegraphics{report_Tomo-editinta_files/figure-latex/unnamed-chunk-1-2.pdf}
